02-antecedentes

### 2024-07-04 02-antecedentes.qmd
Update edit submit

The justification of social inequality based on market-type criteria has been conceptualized as the individuals' adherence to the deservingness of social goods and services (such as health, education, and pensions) based on prices and individuals' ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012]. Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegel_beliefs_1987], highlights that people's explanations and justifications of social inequality are closely tied to their judgments of deservingness. The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations. While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999 ; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations [@lindh_bringing_2023] and even less about children's acceptance or rejection of these market principles. This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.

>>>>

The justification of social inequality based on market-type criteria has been conceptualized as the individuals' adherence to the deservingness of social goods and services (such as health, education, and pensions) based on prices and individuals' ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012]. Research on social stratification beliefs, which explore individual perceptions of who deserves what and why [@kluegel_beliefs_1987], highlights that people's explanations and justifications of social inequality are closely tied to their judgments of deservingness. The influence of ideologies [@wegener_dominant_1995] and cultural schemas [@homan_being_2017] is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations. While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature [@castillo_legitimacy_2011; Evans et al., 2010; @jasso_how_1999; @shariff_income_2016], there has been less examination of public beliefs about which life domains should be governed by market relations [@lindh_bringing_2023] and even less about children's acceptance or rejection of these market principles. This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades [@centeno_arc_2012; @harvey_breve_2015], affecting areas such as pensions, health services, and education.
